MinCall --- MinION od raw signala do genoma

### 2017-05-07 paper/main.tex
Update

\documentclass[times, utf8, seminar, numeric]{fer}
\usepackage{booktabs}
\usepackage{mathtools, amsmath,amsfonts,amssymb, amsthm}
\usepackage{acronym}
\usepackage{geometry}
\usepackage{hyperref}
\usepackage{float}
\usepackage{graphicx}
\graphicspath{ {Figures/} }

@@

\begin{sazetak}

\kljucnerijeci{Basecaller, MinION, sekvenciniranje sljedece generacije}
\end{sazetak}

\begin{abstract}
    The MinION device by Oxford Nanopore is the first portable sequencing device. MinION is able to produce very long reads (reads over 100 kBp were reported), however it suffers from high sequencing error rate. In this paper, we show novel model reducing the error rate can be reduced by improving the base calling process on 1D reads using gated residual convolutional neural networks (CNN) combined with Connectionist Temporal Classification (CTC) learning approach.

@@

A MinION device typically yields reads several thousand bases long; reads as long as 100,000 bp have been reported. 1D read have usually error rate about   30\%. After parameter training, base calling can be performed by running the Viterbi algorithm, which will result in the sequence of states with the highest likelihood.

@@

Earlier models were (Hidden markov model) HMM based where hidden state modeled DNA sequence of length 6 (6-mer) in the nanopore. Pore models were used in computing emission probabilities.~\cite{loman2015complete,schreiber2015analysis,szalay2015novo,timp2012dna} and the recest open source HMM based basecaller Nanocall~\cite{david2016nanocall}.

@@

Dataset were obtained from (Ask Marko!). In this research models were trainined on ecoli K-12 strand. (Should I cite someone for strand). Data was split into train and test subset, such that aligned reads map to different reference genome parts. For initial model data bootstrapping metrichorn was used, version (Ask Someone?).

@@

We're using basecalled knowledge which tells us on each raw read part which 6-mer were currently in the nanopore. Using this data, we get the intermediate target sequence for each block. To correc for model errors, we use aligned information, that is reference genome and alignment data(cigar string) to correct those information, yielding finalized target sequence. For further performace we skip first and last traning block, since most error aggregate on edges.

Adjacent matching bases are spearated with surogate one, for example: AAA -> AA'A for reasong described in following section.

\section{Residual arhitecture}

@@

TODO.

@@

Unlike Nanonet which uses custom OpenCL kernels, or Albacore -- a novel ONT basecaller as of May 2017 lacking GPU support, this work used world class computational framework tensorflow with highly optimized kernels and large development community. Therefore resulting paper's effect is showcasing Residual CNN approach or pure CNN approach with CTC loss is marginally better than already established basecaller, and providing code in contemporary framework.

\bibliographystyle{fer}
\bibliography{refs}
% \bibliographystyle{abbrv}

>>>>

%!TeX spellcheck = en-US
\documentclass[times, utf8, seminar, numeric]{fer}
\usepackage{booktabs}
\usepackage{mathtools, amsmath,amsfonts,amssymb, amsthm}
\usepackage{acronym}
\usepackage{geometry}
\usepackage{hyperref}
\usepackage{float}
\usepackage{graphicx}
\graphicspath{ {Figures/} }

@@

\begin{sazetak}
MinION je citac DNA najnovije generacije. Oxford Nanopore Technologies ga je u svrhu prvog portabilnog DNA sekvencora. Proizvodi vrlo duga ocitanja (duljine i preko 100 kBp), doduse pod cijelu preciznosti. U ovom radu prezentiramo altrenativnu metodu ocitanja DNA koristenjem convolucijskim neuronskih mreza (CNN) umjesto dosadasnjeg pristupa s HMM ili RNN. Takoder cijeli postupak je istreniran od izvornog signala do ocitanih nukleoida, bez ovisnosti o umjetnoj segregaciji na dogadaje.

\kljucnerijeci{Basecaller, MinION, R9, CNN, CTC, sekvenciniranje sljedece generacije}
\end{sazetak}

\begin{abstract}
    The MinION device by Oxford Nanopore is the first portable sequencing device. MinION is able to produce very long reads (reads over 100 kBp were reported), however it suffers from high sequencing error rate. In this paper, we show novel model basecaller which basecalles 1D reads using gated residual convolutional neural networks (CNN) combined with Connectionist Temporal Classification (CTC) learning approach. Its performance is on par with state of the art and offers possible future improvement.

@@

A MinION device typically yields reads several thousand bases long; reads as long as 100,000 bp have been reported. 1D read have usually error rate about   10\%. (Check details!!, What is the error rate, how it's defined???) After parameter training, base calling can be performed by running the Viterbi algorithm, which will result in the sequence of states with the highest likelihood.

@@

Earlier models were (Hidden markov model) HMM based where hidden state modeled DNA sequence of length 6 (6-mer) in the nanopore. Pore models were used in computing emission probabilities.~\cite{loman2015complete,schreiber2015analysis,szalay2015novo,timp2012dna} and the recent open source HMM based basecaller Nanocall~\cite{david2016nanocall}.

@@

Dataset were obtained from (Ask somebody, Marko doesn't know!). In this research models were trainined on ecoli K-12 strand. (Should I cite someone for strand). Data was split into train and test subset, such that aligned reads map to different reference genome parts. For initial model data bootstrapping metrichorn was used, version (Ask Someone?).

@@

We're using basecalled knowledge which tells us on each raw read part which 6-mer were currently in the nanopore. Using this data, we get the intermediate target sequence for each block. To correct for model errors, we use aligned information, that is reference genome and alignment data(cigar string) to correct those information, yielding finalized target sequence. For further performace we skip first and last traning block, since most error aggregate on edges.

Adjacent matching bases are spearated with surogate one, for example: AAA -> AA'A for reasong described in following section.
\section{Residual arhitecture}

@@

TODO

See \url{https://github.com/nmiculinic/minion-basecaller/blob/master/notebook/Lambda_nanonet_vs_m270.ipynb for preliminary results.}

@@

Unlike Nanonet which uses custom OpenCL kernels, or Albacore --- a novel ONT basecaller as of May 2017 lacking GPU support, this work used world class computational framework tensorflow with highly optimized kernels and large development community. Therefore resulting paper's effect is showcasing Residual CNN approach or pure CNN approach with CTC loss is marginally better than already established basecaller, and providing code in contemporary framework.

\bibliographystyle{fer}
\bibliography{refs}
